GSE158422: DNA methylation in Lung squamous cell carcinoma - GMEA using fry()

### 2022-10-11 GSE158433/GSE158422.Rmd
getting fry to work

SERIES_MATRIX="GSE158422_series_matrix.txt.gz"
if(!file.exists(SERIES_MATRIX)){
download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE158nnn/GSE158422/matrix/GSE158422_series_matrix.txt.gz",
  destfile=SERIES_MATRIX)
}
gse <- getGEO(filename=SERIES_MATRIX)
baseDir <- "."
sample_metadata <- pData(phenoData(gse))
targets <- sample_metadata
mybase <- unique(gsub("_Red.idat.gz" ,"",  gsub("_Grn.idat.gz", "" ,list.files(".",pattern = "GSM",recursive = TRUE))))
targets$Basename <- mybase
head(targets)
rgSet <- read.metharray.exp(targets = targets)
rgSet

@@

par(mfrow=c(2,1))  
myscree(Mval,main="incl sex chr")
myscree(Mval_flt,main="excl sex chr")

@@

samplesheet<-targets
groups <- factor(samplesheet$case,levels=c("tumor: normal","tumor: tumor"))
sex <- factor(samplesheet$sex,levels=c("0","1"))
Mvals<-Mval
betas<-beta
dim(Mvals)
dim(betas)

GSE158422 <- dm_analysis(samplesheet=samplesheet,
  sex=sex,groups=groups,mx=Mvals,name="GSE158422",
  myann=myann ,beta= betas)

>>>>

SERIES_MATRIX="GSE158422_series_matrix.txt.gz"
if(!file.exists(SERIES_MATRIX)){
download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE158nnn/GSE158422/matrix/GSE158422_series_matrix.txt.gz",
  destfile=SERIES_MATRIX)
}
gse <- getGEO(filename=SERIES_MATRIX)
baseDir <- "IDAT"
sample_metadata <- pData(phenoData(gse))
targets <- sample_metadata
mybase <- unique(gsub("_Red.idat.gz" ,"",  gsub("_Grn.idat.gz", "" ,list.files(".",pattern = "GSM",recursive = TRUE))))
targets$Basename <- mybase
head(targets)
rgSet <- read.metharray.exp(targets = targets)
rgSet

targets$Basename <- gsub("IDAT/","",mybase)
head(targets)

@@

par(mfrow=c(2,1))
myscree(Mval,main="incl sex chr")
myscree(Mval_flt,main="excl sex chr")
par(mfrow=c(1,1))

@@

samplesheet<-targets
groups <- factor(samplesheet$case,levels=c("tumor: normal","tumor: tumor"))
sex <- factor(samplesheet$sex,levels=c("0","1"))
dim(Mval)
dim(beta)

GSE158422 <- dm_analysis(samplesheet=samplesheet,
  sex=sex,groups=groups,mx=Mval,name="GSE158422",
  myann=myann ,beta= beta)

### 2022-10-11 GSE158433/GSE158422_fry.Rmd
getting fry to work

suppressPackageStartupMessages({
  library("plyr")
  library("R.utils")
  library("missMethyl")
  library("limma")
  library("DMRcate")
  library("DMRcatedata")
  library("topconfects")
  library("minfi")
  library("IlluminaHumanMethylation450kmanifest")
  library("RColorBrewer")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("GEOquery")
  library("eulerr")
  library("plyr")
  library("gplots")
  library("reshape2")
  library("forestplot")
  library("beeswarm")
  library("RCircos")
  library("qqman")
  library("ENmix")

@@

## Differential analysis

Paired normal versus cancer tissue.

```{r,dm}

samplesheet<-targets
groups <- factor(samplesheet$case,levels=c("tumor: normal","tumor: tumor"))
sex <- factor(samplesheet$sex,levels=c("0","1"))
Mvals<-Mval
betas<-beta
dim(Mvals)
dim(betas)

GSE158422 <- dm_analysis(samplesheet=samplesheet,
  sex=sex,groups=groups,mx=Mvals,name="GSE158422",
  myann=myann ,beta= betas)

mx=Mvals
name="GSE158422"
beta=betas
design <- model.matrix(~ sex + groups)
mxs <- mx[,which( colnames(mx) %in% samplesheet$Basename )]
fit.reduced <- lmFit(mxs,design)
fit.reduced <- eBayes(fit.reduced)
summary(decideTests(fit.reduced))

>>>>

suppressPackageStartupMessages({
  library("plyr")
  library("R.utils")
  library("missMethyl")
  library("limma")
  library("DMRcate")
  library("DMRcatedata")
  library("topconfects")
  library("minfi")
  library("IlluminaHumanMethylation450kmanifest")
  library("RColorBrewer")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("GEOquery")
  library("eulerr")
  library("plyr")
  library("gplots")
  library("reshape2")
  library("forestplot")
  library("beeswarm")
  library("RCircos")
  library("qqman")
  library("ENmix")

@@

## Design matrix

```{r,design}

samplesheet<-targets
groups <- factor(samplesheet$case,levels=c("tumor: normal","tumor: tumor"))
sex <- factor(samplesheet$sex,levels=c("0","1"))
design <- model.matrix(~ sex + groups)
mxs <- Mval
